---
title: "COVID19 Analysis"
author: "Jingyang(Judy) Zhang"
date: "7/13/2020"
output: html_document
---

# Load Packages
```{r}
library(tidyverse)
library(readr)
library(janitor)
library(wbstats)
```

# Load Source Files
```{r}
source("map_covid.R")
```


# Load data

## BORDERING COUNTRY DATA ##
```{r}
borders_url <- url("https://raw.githubusercontent.com/geodatasource/country-borders/master/GEODATASOURCE-COUNTRY-BORDERS.CSV")

borders <- read_csv(borders_url) %>% clean_names()
```


## COVID19 DATA ##
```{r}
date_str <- "05-22-2020"

covid19 <- getDat(date_str) %>% clean_coviddat()
```


## Data Manipulation ##
Match country names in borders to country names in covid19
```{r}
borders_countries <- tolower(unique(borders$country_name))

covid19_countries <- covid19$country_region

sort(setdiff(borders_countries, covid19_countries))

# aland islands: not in covid19
# american samoa: not in covid19
# anguilla: not in covid19
# antarctica: not in covid19
# aruba: not in covid19
# bermuda: not in covid19
# bolivia (plurinational state of): bolivia in covid19
borders$country_name[tolower(borders$country_name) == "bolivia (plurinational state of)"] <- "bolivia"

borders$country_border_name[tolower(borders$country_border_name) == "bolivia (plurinational state of)"] <- "bolivia"

# bonaire, sint eustatius and saba: not in covid19
# bouvet island: not in covid19
# british indian ocean territory: not in covid19
# brunei darussalam: brunei in covid19
borders$country_name[tolower(borders$country_name) == "brunei darussalam"] <- "brunei"
borders$country_border_name[tolower(borders$country_border_name) == "brunei darussalam"] <- "brunei"

# cayman islands: not in covid19
# christmas island: not in covid19
# cocos (keeling) islands: not in covid19
# congo: congo (brazzaville) in covid19
borders$country_name[tolower(borders$country_name) == "congo"] <- "congo (brazzaville)"

borders$country_border_name[tolower(borders$country_border_name) == "congo"] <- "congo (brazzaville)"

# congo (the democratic republic of the): congo (kinshasa) in covid19
borders$country_name[tolower(borders$country_name) == "congo (the democratic republic of the)"] <- "congo (kinshasa)"

borders$country_border_name[tolower(borders$country_border_name) == "congo (the democratic republic of the)"] <- "congo (kinshasa)"
# cook islands: not in covid19
# cote d’ivoire: cote d'ivoire in covid19
borders$country_name[tolower(borders$country_name) == "cote d’ivoire"] <- "cote d'ivoire"

borders$country_border_name[tolower(borders$country_border_name) == "cote d’ivoire"] <- "cote d'ivoire"

# curacao: not in covid19
# falkland islands (malvinas): not in covid19
# faroe islands: not in covid19
# french guiana: not in covid19
# french polynesia: not in covid19
# french southern territories: not in covid19
# gambia (the): gambia in covid19
borders$country_name[tolower(borders$country_name) == "gambia (the)"] <- "gambia"

borders$country_border_name[tolower(borders$country_border_name) == "gambia (the)"] <- "gambia"
# gibraltar: not in covid19
# greenland: not in covid19
# guadeloupe: not in covid19
# guam: not in covid19
# guernsey: not in covid19
# heard island and mcdonald islands: not in covid19
# hong kong: not in covid19
# iran (islamic republic of): iran in covid19
borders$country_name[tolower(borders$country_name) == "iran (islamic republic of)"] <- "iran"

borders$country_border_name[tolower(borders$country_border_name) == "iran (islamic republic of)"] <- "iran"
# isle of man: not in covid19
# jersey: not in covid19
# kiribati: not in covid19
# korea (democratic people's republic of): korea, south in covid19
borders$country_name[tolower(borders$country_name) == "korea (democratic people's republic of)"] <- "korea, south"

borders$country_border_name[tolower(borders$country_border_name) == "korea (democratic people's republic of)"] <- "korea, south"
# korea (the republic of): not in covid19
# lao people's democratic republic: laos in covid19
borders$country_name[tolower(borders$country_name) == "lao people's democratic republic"] <- "laos"

borders$country_border_name[tolower(borders$country_border_name) == "lao people's democratic republic"] <- "laos"
# macao: not in covid19
# marshall islands: not in covid19
# martinique: not in covid19
# mayotte: not in covid19
# micronesia (federated states of): not in covid19
# moldova (the republic of): moldova in covid19
borders$country_name[tolower(borders$country_name) == "moldova (the republic of)"] <- "moldova"

borders$country_border_name[tolower(borders$country_border_name) == "moldova (the republic of)"] <- "moldova"
# montserrat: not in covid19
# myanmar: burma in covid19
borders$country_name[tolower(borders$country_name) == "myanmar"] <- "burma"

borders$country_border_name[tolower(borders$country_border_name) == "myanmar"] <- "burma"
# nauru: not in covid19
# new caledonia: not in covid19
# niue: not in covid19
# norfolk island: not in covid19
# northern mariana islands: not in covid19
# palau: not in covid19
# palestine, state of: not in covid19
# pitcairn: not in covid19
# puerto rico: not in covid19
# reunion: not in covid19
# russian federation: russia in covid19
borders$country_name[tolower(borders$country_name) == "russian federation"] <- "russia"

borders$country_border_name[tolower(borders$country_border_name) == "russian federation"] <- "russia"
# saint barthelemy: not in covid19
# saint helena,\"ascension and tristan da cunha: not in covid19
# saint martin (french part): not in covid19
# saint pierre and miquelon: not in covid19
# samoa:
# sint maarten (dutch part): not in covid19
# solomon islands: not in covid19
# south georgia and the south sandwich islands: not in covid19
# svalbard and jan mayen: not in covid19
# syrian arab republic: syria in covid19
borders$country_name[tolower(borders$country_name) == "syrian arab republic"] <- "syria"

borders$country_border_name[tolower(borders$country_border_name) == "syrian arab republic"] <- "syria"
# taiwan (province of china): not in covid19
# tanzania (the united republic of): tanzania in covid19
borders$country_name[tolower(borders$country_name) == "tanzania (the united republic of)"] <- "tanzania"

borders$country_border_name[tolower(borders$country_border_name) == "tanzania (the united republic of)"] <- "tanzania"
# tokelau: not in covid19
# tonga: not in covid19
# turkmenistan: not in covid19
# turks and caicos islands: not in covid19
# tuvalu: not in covid19
# united kingdom of great britain and northern ireland: not in covid19
# united states minor outlying islands: not in covid19
# united states of america: us in covid19
borders$country_name[tolower(borders$country_name) == "united states of america"] <- "us"

borders$country_border_name[tolower(borders$country_border_name) == "united states of america"] <- "us"
# vanuatu: not in covid19
# venezuela (bolivarian republic of): venezuela in covid19
borders$country_name[tolower(borders$country_name) == "venezuela (bolivarian republic of)"] <- "venezuela"

borders$country_border_name[tolower(borders$country_border_name) == "venezuela (bolivarian republic of)"] <- "venezuela"
# viet nam: vietname in covid19
borders$country_name[tolower(borders$country_name) == "viet nam"] <- "vietnam"

borders$country_border_name[tolower(borders$country_border_name) == "viet nam"] <- "vietnam"
# virgin islands (british): not in covid19
# virgin islands (u.s.): not in covid19
# wallis and futuna: not in covid19




```




## Create Connectivity Matrix (Weight Matrix) ##

Use binary connectivity matrix (i.e. adjacency weights) with $w_{ij} = 1$ if country $i$ and country $j$ share any portion of their borders with another.

We will build the matrix using the `borders` dataset.

## Function: get_borders
## Gets border countries of a given country
## Arg: 
## borders: dataset containing border information
## country: country of interest
## Return: list of countries that share border with the given country
```{r}
get_borders <- function(borders, country){
  border_countries <- borders %>% filter(tolower(country_name) == tolower(country))
  
  border_countries <- tolower(border_countries$country_border_name)
  
  return(border_countries)
}

```






## Function: get_connect
## Creates a connect matrix using information from the borders dataset
## Arg: 
## borders: dataset containing border information
## Return: a connectivity matrix 

```{r}
get_connect <- function(borders){
  # Get list of unique countries 
  countries <- tolower(unique(borders$country_name))
  # Create an empty connect matrix
  connect_matrx <- matrix(data = NA, nrow = length(countries), ncol = length(countries))
  colnames(connect_matrx) <- countries
  rownames(connect_matrx) <- countries
  
  # Get border information of every country in the countries list
  for(i in 1 : length(countries)){
    country <- countries[i]
    border_countries <- get_borders(borders, country)
   
    if(is.na(border_countries)){
      next
    }
    else{
      
    if(length(border_countries) > 1){
      for(j in 1 : length(border_countries)){
        
        connect_matrx[border_countries[j], country] <- 1
      }
    }
    else{
     
      connect_matrx[border_countries, country] <- 1
    }
      
    }
    
  }
  
  # Replace all NA with 0s
  connect_matrx[is.na(connect_matrx)] <- 0
  
  return(connect_matrx)
  
  
}



# Get connectivity matrix
connect_matrx <- get_connect(borders)

```



## Function: value_lookup
## Look up value of confirmed/recovered/deaths cases of a given country
## Arg: 
## country: country of interest
## covid19: covid19 dataset
## type: which type of value to draw, confirmed, recovered, deaths
## Return: a number of value of confirmed/recovered/deaths cases of the given county
```{r}
value_lookup <- function(covid19, country, type){
  country_value <- covid19 %>% 
    filter(tolower(country_region) == tolower(country))
  
  case_value <- case_when(
    type == "confirmed" ~ country_value$total_confirmed,
    type == "recovered" ~ country_value$total_recovered,
    type == "deaths" ~ country_value$total_deaths
  )
  
  if(length(case_value) == 0){
    case_value <- 0
  }
  
  return(case_value)
  
}
```


## Calculate Variance of Attribute Values ##

```{r}
attr_vals <- covid19$total_confirmed

var_attr <- var(attr_vals)

## HAND CALCULATION
#n <- length(attr_vals)
#mean_attr <- mean(attr_vals)
#s2_attr <- (1 / (n - 1)) * sum((attr_vals - mean_attr)^2)
```


## Calculate Mean of Attribute Values ##
```{r}
n <- length(attr_vals)
mean_attr <- mean(attr_vals)
```


## Calculate Sum of Weights, $w_{..}$ ##

```{r}
sum_weights <- sum(connect_matrx)
```




### Caculate Moran's I ###
$I = \frac{n}{(n-1)S^{2}w_{..}}\sum_{i=1}^{n}\sum_{j=1}^{n}[(Z(s_{i})-\bar{Z})(Z(s_{j})-\bar{Z})]$
```{r}
type <- "confirmed"
weighted <- c()
for(i in 1 : nrow(connect_matrx)){
  for(j in 1: nrow(connect_matrx)){
    si <- rownames(connect_matrx)[i]
    sj <- rownames(connect_matrx)[j]
    zsi <- value_lookup(covid19, si, type)
    zsj <- value_lookup(covid19, sj, type)
  
    value <- connect_matrx[i, j] * (zsi - mean_attr) * (zsj - mean_attr)
    
    weighted[i] <- value
    
  }
}

I <- (n / ((n - 1) * var_attr * sum_weights)) * sum(weighted)

```



Standardize I Test Statistic:

```{r}
EI <- (-1 / (n - 1))

```


$S_{1} = \frac{1}{2}\sum_{i=1}^{n}\sum_{j=1}^{n}(w_{ij} + w_{ji})^{2}$

```{r}
#### S1 #####
sum_ij <- c()
for(i in 1 : nrow(connect_matrx)){
  for(j in 1 : nrow(connect_matrx)){
    
    value <- (connect_matrx[i, j] + connect_matrx[j, i])^2
    
    sum_ij[i] <- value
  }
}

S1 <- (1 / 2) * sum(sum_ij)
```

#### S2 #####

$S_{2} = \sum_{i=1}^{n}(\sum_{j = 1}^{n}w_{ij} + \sum_{j = 1}^{n}w_{ji})^{2}$

```{r}
sum_i <- c()

for(i in 1 : nrow(connect_matrx)){
  sum_wij <- sum(connect_matrx[i,])
  sum_wji <- sum(connect_matrx[,j])
  
  value <- (sum_wij + sum_wji)^2
  sum_i[i] <- value
  
}

S2 <- sum(sum_i)




```



#### S3 #####


$S_{3} = \frac{n^{-1}\sum_{i = 1}^{n}(Z(s_{i})-\bar{Z})^{4}}{(n^{-1}\sum_{i=1}^{n}(Z(s_{i})-\bar{Z})^{2})^{2}}$


```{r}

sum_num <- c()

for(i in 1 : nrow(connect_matrx)){
    si <- rownames(connect_matrx)[i]

    zsi <- value_lookup(covid19, si, type)
    
    value <- (zsi - mean_attr)^4
  
    sum_num[i] <- value
  
}



sum_denom <- c()

for(i in 1 : nrow(connect_matrx)){
    si <- rownames(connect_matrx)[i]

    zsi <- value_lookup(covid19, si, type)
    
    value <- (zsi - mean_attr)^2
  
    sum_denom[i] <- value
  
}


S3 <- ((1 / n) * sum(sum_num)) / ((1 / n) * sum(sum_denom))^2


```




#### S4 #####


$S_{4} = (n^{2} - 3n + 3)S_{1} - nS_{2} + 3w_{..}$


```{r}

S4 <- (n^2 - 3 * n + 3) * S1 - n * S2 + 3 * sum_weights^2


```




#### S5 #####


$S_{5} = (n^{2} - n)S_{1} - 2nS_{2} + 6w_{..}$


```{r}

S5 <- (n^2 - n) * S1 - 2 * n * S2 + 6 * sum_weights^2


```


Calculate Variance of Moran's I under null distribution

Var(I)$=\frac{nS_{4}-S_{3}S_{5}}{(n-1)(n-2)(n-3)w_{..}^{2}} - (E(I))^{2}$
```{r}


varI <- (n * S4 - S3 * S5) / (n - 1) * (n - 2) * (n - 3) * sum_weights^2 - EI^2

```




Calculate Z-score and p-value

```{r}
z_score <- (I - EI) / varI

p_value <- pnorm(abs(z_score))

p_value

```

