---
title: "COVID19 Analysis"
author: "Jingyang(Judy) Zhang"
date: "7/13/2020"
output: html_document
---

# Load Packages
```{r}
library(tidyverse)
library(readr)
library(janitor)
```

# Load Source Files
```{r}
source("map_covid.R")
```


# Load data

## BORDERING COUNTRY DATA ##
```{r}
borders_url <- url("https://raw.githubusercontent.com/geodatasource/country-borders/master/GEODATASOURCE-COUNTRY-BORDERS.CSV")

borders <- read_csv(borders_url) %>% clean_names()
```


## COVID19 DATA ##
```{r}
date_str <- "05-22-2020"

covid19 <- getDat(date_str) %>% clean_coviddat()
```


## Data Manipulation ##
Match country names in borders to country names in covid19
```{r}
borders_countries <- tolower(unique(borders$country_name))

covid19_countries <- covid19$country_region

sort(setdiff(borders_countries, covid19_countries))

# aland islands: not in covid19
# american samoa: not in covid19
# anguilla: not in covid19
# antarctica: not in covid19
# aruba: not in covid19
# bermuda: not in covid19
# bolivia (plurinational state of): bolivia in covid19
borders$country_name[tolower(borders$country_name) == "bolivia (plurinational state of)"] <- "bolivia"

# bonaire, sint eustatius and saba: not in covid19
# bouvet island: not in covid19
# british indian ocean territory: not in covid19
# brunei darussalam: brunei in covid19
borders$country_name[tolower(borders$country_name) == "brunei darussalam"] <- "brunei"

# cayman islands: not in covid19
# christmas island: not in covid19
# cocos (keeling) islands: not in covid19
# congo: congo (brazzaville) in covid19
borders$country_name[tolower(borders$country_name) == "congo"] <- "congo (brazzaville)"

# congo (the democratic republic of the): congo (kinshasa) in covid19
borders$country_name[tolower(borders$country_name) == "congo (the democratic republic of the)"] <- "congo (kinshasa)"

# cook islands: not in covid19
# cote d’ivoire: cote d'ivoire in covid19
borders$country_name[tolower(borders$country_name) == "cote d’ivoire"] <- "cote d'ivoire"

# curacao: not in covid19
# falkland islands (malvinas): not in covid19
# faroe islands: not in covid19
# french guiana: not in covid19
# french polynesia: not in covid19
# french southern territories: not in covid19
# gambia (the): gambia in covid19
borders$country_name[tolower(borders$country_name) == "gambia (the)"] <- "gambia"
# gibraltar: not in covid19
# greenland: not in covid19
# guadeloupe: not in covid19
# guam: not in covid19
# guernsey: not in covid19
# heard island and mcdonald islands: not in covid19
# hong kong: not in covid19
# iran (islamic republic of): iran in covid19
borders$country_name[tolower(borders$country_name) == "iran (islamic republic of)"] <- "iran"
# isle of man: not in covid19
# jersey: not in covid19
# kiribati: not in covid19
# korea (democratic people's republic of): korea, south in covid19
borders$country_name[tolower(borders$country_name) == "korea (democratic people's republic of)"] <- "korea, south"
# korea (the republic of): not in covid19
# lao people's democratic republic: laos in covid19
borders$country_name[tolower(borders$country_name) == "lao people's democratic republic"] <- "laos"
# macao: not in covid19
# marshall islands: not in covid19
# martinique: not in covid19
# mayotte:
# micronesia (federated states of):
# moldova (the republic of): 
# montserrat:
# myanmar:
# nauru:
# new caledonia:
# niue:
# norfolk island:
# northern mariana islands:
# palau: 
# palestine, state of:
# pitcairn:
# puerto rico:
# reunion:
# russian federation:
# saint barthelemy:
# saint helena,\"ascension and tristan da cunha:
# saint martin (french part):
# saint pierre and miquelon:
# samoa:
# sint maarten (dutch part):
# solomon islands:
# south georgia and the south sandwich islands:
# svalbard and jan mayen:
# syrian arab republic:
# taiwan (province of china): not in covid19
# tanzania (the united republic of): tanzania in covid19
borders$country_name[tolower(borders$country_name) == "tanzania (the united republic of)"] <- "tanzania"
# tokelau: not in covid19
# tonga: not in covid19
# turkmenistan: not in covid19
# turks and caicos islands: not in covid19
# tuvalu: not in covid19
# united kingdom of great britain and northern ireland: not in covid19
# united states minor outlying islands: not in covid19
# united states of america: us in covid19
borders$country_name[tolower(borders$country_name) == "united states of america"] <- "us"
# vanuatu: not in covid19
# venezuela (bolivarian republic of): venezuela in covid19
borders$country_name[tolower(borders$country_name) == "venezuela (bolivarian republic of)"] <- "venezuela"
# viet nam: vietname in covid19
borders$country_name[tolower(borders$country_name) == "viet nam"] <- "vietnam"
# virgin islands (british): not in covid19
# virgin islands (u.s.): not in covid19
# wallis and futuna: not in covid19



```




## Create Connectivity Matrix (Weight Matrix) ##

Use binary connectivity matrix (i.e. adjacency weights) with $w_{ij} = 1$ if country $i$ and country $j$ share any portion of their borders with another.

We will build the matrix using the `borders` dataset.

## Function: get_borders
## Gets border countries of a given country
## Arg: 
## borders: dataset containing border information
## country: country of interest
## Return: list of countries that share border with the given country
```{r}
get_borders <- function(borders, country){
  border_countries <- borders %>% filter(country_name == country)
  
  border_countries <- border_countries$country_border_name
  
  return(border_countries)
}

```


```{r}
countries <- unique(borders$country_name)

connect_matrx <- matrix(data = NA, nrow = length(countries), ncol = length(countries))

colnames(connect_matrx) <- countries

for(i in 1 : nrow(connect_matrx)){
  for(j in 1 : ncol(connect_matrx)){
    connect_matrx[i, j] <- countries[i]
  }
}




```



## Function: get_connect
## Creates a connect matrix using information from the borders dataset
## Arg: 
## borders: dataset containing border information
## Return: a connectivity matrix 

```{r}
get_connect <- function(borders){
  # Get list of unique countries 
  countries <- unique(borders$country_name)
  # Create an empty connect matrix
  connect_matrx <- matrix(data = NA, nrow = length(countries), ncol = length(countries))
  colnames(connect_matrx) <- countries
  rownames(connect_matrx) <- countries
  
  # Get border information of every country in the countries list
  for(i in 1 : length(countries)){
    country <- countries[i]
    border_countries <- get_borders(borders, country)
    if(is.na(border_countries)){
      next
    }
    else{
      
    if(length(border_countries) > 1){
      for(j in 1 : length(border_countries)){
        connect_matrx[border_countries[j], country] <- 1
      }
    }
    else{
      connect_matrx[border_countries, country] <- 1
    }
      
    }
    
  }
  
  # Replace all NA with 0s
  connect_matrx[is.na(connect_matrx)] <- 0
  
  return(connect_matrx)
  
  
}



# Get connectivity matrix
connect_matrx <- get_connect(borders)

```



## Calculate Variance of Attribute Values ##

```{r}
attr_vals <- covid19$total_confirmed

var_attr <- var(attr_vals)

## HAND CALCULATION
#n <- length(attr_vals)
#mean_attr <- mean(attr_vals)
#s2_attr <- (1 / (n - 1)) * sum((attr_vals - mean_attr)^2)
```


## Calculate Mean of Attribute Values ##
```{r}
n <- length(attr_vals)
mean_attr <- mean(attr_vals)
```


## Calculate Sum of Weights, $w_{..}$ ##

```{r}
sum_weights <- sum(connect_matrx)
```




### Caculate Moran's I ###
```{r}

for(i in 1 : nrow(connect_matrx)){
  for(j in 1: ncol(connect_matrx)){
    
  }
}

```

